---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# scotgov

[![Travis-CI Build Status](https://travis-ci.org/jsphdms/scotgov.svg?branch=master)](https://travis-ci.org/jsphdms/scotgov)

:construction: :construction: :construction: **Package under construction - watch this space for updates** :construction: :construction: :construction:

Use scotgov to download data from [statistics.gov.scot](http://statistics.gov.scot/home) with a single line of R code. scotgov removes the need to write SPARQL code; you simply need the URI of a dataset. scotgov can be used interactively, or as part of a [reproducible analytical pipeline](https://ukgovdatascience.github.io/rap_companion/).

## Installation

Install scotgov from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("jsphdms/scotgov")
```

If the above does not work, you can install from source:

1. Go to the scotgov [repository](https://github.com/jsphdms/scotgov) on GitHub
1. Click **Clone or download** then **Download ZIP**
1. Save the file locally (e.g. your H drive) and Unzip
1. Install with install.packages()

```{r source-installation, eval = FALSE}
install.packages("your/directory/scotgov", repos = NULL,
                 type="source", lib = "your/R/package/library/directory")
```

Usage
-----

Find the name of the dataset you want using `list_sg_dataset()`

Learn more in `vignette("scotgov")` or `?scotgov_get`.

Use `list_sg_dataset()` to find all datasets currently loaded onto statistics.gov.scot

Use `scotgov_get()` to get the data from a dataset in statistics.gov.scot

Use `get_structure()` to find all the full sets of categories and values for a particular dataset (helpful for creating new filters for `scotgov_get`!)

Use `get_dataset_query()` to produce the SPARQL query used by `scotgov_get()`.

Examples
---------

Get a dataframe of all datasets on statistics.gov.scot, their uri, and publisher
`all_scotgov_datasets <- list_sg_datasets()`

Discover the structure of the dataset on homelessness applications - so we can use this in a later filter
`our_structure <- get_structure("homelessness-applications")`

After viewing our_structure - we decide we only want the data for "all applications" and for the periods "2015/2016" and "2016/2017", so we add these to the filter. 

``` r
filtered_data <- scotgov_get("homelessness-applications",
                              refPeriod = c("2015/2016", "2016/2017"),
                              applicationType = "All applications")
```
                                                 
Future development
------------------

This package is under active development, so any further functionality will be mentioned here when it's ready. If something important is missing, feel free to contact the contributors or [add a new issue](https://github.com/jsphdms/scotgov/issues).

Since this package is under active development, breaking changes may be necessary. We will make it clear once the package is reasonably stable.
