---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# scotgov

[![Travis-CI Build Status](https://travis-ci.org/jsphdms/scotgov.svg?branch=master)](https://travis-ci.org/jsphdms/scotgov)

[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)

:construction: :construction: :construction: **Package under construction - watch this space for updates** :construction: :construction: :construction:

Use scotgov to download data from [statistics.gov.scot](http://statistics.gov.scot/home) with a single line of R code. scotgov removes the need to write SPARQL code; you simply need the URI of a dataset. scotgov can be used interactively, or as part of a [reproducible analytical pipeline](https://ukgovdatascience.github.io/rap_companion/).

## Installation

Install from GitHub:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("jsphdms/scotgov")
```

Or you can install from source:

1. Go to the scotgov [repository](https://github.com/jsphdms/scotgov) on GitHub
1. Click **Clone or download** then **Download ZIP**
1. Save the file locally (e.g. your H drive)
1. Unzip the file
1. Install with `install.packages()`

```{r source-installation, eval = FALSE}
install.packages("your/directory/scotgov-master/scotgov-master", repos = NULL,
                 type="source", lib = "your/R/package/library/directory")
```

Having trouble installing from source? Check you followed the steps above precisely; a small error can cause the install to fail. The best approach is to copy the code above into a text editor and carefully update the two directories. In particular, notice the repetition of **scotgov-master**.

## Usage

Learn more in `vignette("scotgov")` or `?scotgov_get`.

## Future development

Currently `scotgov_get()` is the only function available. This package is under active development, so any further functionality will be mentioned here when it's ready. If something important is missing, feel free to contact the contributors or [add a new issue](https://github.com/jsphdms/scotgov/issues).

Since this package is under active development, breaking changes may be necessary. We will make it clear once the package is reasonably stable.
