---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# scotgov

Use scotgov to download data from [statistics.gov.scot](http://statistics.gov.scot/home) with a single line of R code. scotgov removes the need to write SPARQL code; you simply need the URI of a dataset. scotgov can be used interactively, or as part of a [reproducible analytical pipeline](https://ukgovdatascience.github.io/rap_companion/).

## Installation

Install scotgov from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("jsphdms/scotgov")
```

If the above does not work, you can install from source:

1. Go to the scotgov [repository](https://github.com/jsphdms/scotgov) on GitHub
1. Click **Clone or download** then **Download ZIP**
1. Save the file locally (e.g. your H drive) and Unzip
1. Install with install.packages()

```{r source-installation, eval = FALSE}
install.packages("your/directory/scotgov", repos = NULL,
                 type="source", lib = "your/R/package/library/directory")
```

## Example

You will need the URI for your dataset. Find this on the [statistics.gov.scot](http://statistics.gov.scot/home) web page for your dataset (in the API tab):

```{r example, message = FALSE}
library(scotgov)

household_size <- scotgov_get("http://statistics.gov.scot/data/average-household-size")
head(household_size)
```
