---
title: "Introduction to scotgov package"
author: "Joseph Adams"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{scotgov introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(scotgov)
```


## Introduction

Use _scotgov_ to download data from [statistics.gov.scot](https://statistics.gov.scot) with a single line of R code. _scotgov_ removes the need to write SPARQL code; you simply need the URI of a dataset. _scotgov_ can be used interactively, or as part of a [reproducible analytical pipeline](https://ukgovdatascience.github.io/rap_companion/).


## Usage

You can download an entire dataset, or filter by date and/or geography. We recommend filtering large datasets. If you require a full download of a large dataset, you may need to contact [statistics.gov.scot](https://statistics.gov.scot).  
<br/>

You will always need the last part of the URI for your dataset. Find this on the [statistics.gov.scot](https://statistics.gov.scot) web page for your dataset (in the API tab). For example, the full URI for Average Household Size is [Average Household Size](https://statistics.gov.scot/resource?uri=http%3A%2F%2Fstatistics.gov.scot%2Fdata%2Faverage-household-size):  


`http://statistics.gov.scot/data/average-household-size`

you just need the last part:  

`average-household-size`


### Without filtering

Once you have the last part of the URI, set this as the `dataset` parameter for `scotgov_get()`:  

```{r}
household_size <- scotgov_get(dataset = "average-household-size")
head(household_size)
```

### With filtering

Filtering is useful for large data sets. The filtering parameters for `scotgov_get()` are:

-   `start_date`
-   `end_date`
-   `geography`

##### Filter by date

Use either `start_date` **OR** `end_date` to filter datapoints before or after a certain date:

``` {r}
household_size_2010_onwards <- scotgov_get(dataset = "average-household-size",
                                           start_date = 2010)
head(household_size_2010_onwards)

```

Use `start_date` **AND** `end_date` to filter datapoints within a certain timeframe.

##### Filter by geography


Specify a single geography using an S code:

``` {r}
household_size_S12000039 <- scotgov_get(dataset = "average-household-size",
                                           geography = "S12000039")
head(household_size_S12000039)
```

-------


```{r}
sessionInfo()
```
